---
title: "Network Construction and Simulation"
output:
  html_document:
    toc: true
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```


## Loading Tf-target database
Many times a context specific network is needed to explain the experimental 
data. One can start with a known literature based network interactions or 
regulons obtained from other data or methods. A regulon is typically a TF and 
its targets possibly with a weight metric. Such regulons may obtained from
activity calculation methods like SCENIC etc. or from other datasets like
ChIP seq or ATAC seq.

```{r regulons}
# Laoding data from a known database.
# network <- readRDS("database.rds") 
# Here network is Source,Target two column dataframe.

# If the network is in form of a list
listNetwork <- readRDS("database.rds") # like hDB_v2

network <- data.frame(Source = rep(names(lapply(listNetwork, function(x)names(x))), times = unlist(lapply(listNetwork, function(x)length(x)))),
                      Target = unlist(hDB_v2))

# From SCENIC results
# regs <- read_json("regulon.json", simplify = T)
# network <- data.frame(Source = character(), Target = character())
# for(src in seq_along(regs)){
#   targets <- unlist(regs[src])
#   networkTmp <- data.frame(Source = rep(names(regs)[src],length(targets)), Target = targets)
#   network <- rbind(network, networkTmp)
# }

```

## Filter the network based on core transcription factors
```{r}

# Core transcription factors around which to built the network. Only the
# TFs which are regulator or target of these coreTfs in the database
# will be included in the network.

coreTf <- readRDS(("coreTf.rds"))

network <- network[union(which(network$Source %in% coreTf),
                                 which(network$Source %in% coreTf)),]
  
```
## Add correlation to the network

```{r}
      expData <- readRDS("expData.rds")
      networkGenes <- union(network$Source,network$Target)
      expData <- t(expData[networkGenes,])
      actCor <- cor(expData, method = "s")
      type <- integer(length = length(network$Source))
      for(int in seq_along(network$Source)){
        type[int] <- actCor[as.character(network$Source[int]), as.character(network$Target[int])]
      }
      # Correlation
      network$Cor <- type
      # Assign interaction sign based on correlation
      type[type>0] <- 1
      type[type<=0] <- 2
      network$Type <- type
```

## Add mutual information to the network.

Function to calculate mutual information.
```{r}

library(infotheo)
#' @param data expression matrix. Genes in rows and samples/cells in columns
#' @param nbins number of bins
#' @value A mutual information matrix
calculateMI <- function(data = actMat, nbins=16){
  data <- t(data)
  nGenes <- dim(data)[2]
  miMat <- matrix(0,nrow = nGenes,ncol = nGenes)
  geneNames <- colnames(data)
  rownames(miMat) <- geneNames
  colnames(miMat) <- geneNames
  data <- infotheo::discretize(data)
  #i=1
  for(i in 1:(nGenes-1))
  {
    for(j in (i+1):(nGenes))
    {
      gene1 <- geneNames[i]
      gene2 <- geneNames[j]
      miMat[gene1, gene2] <- infotheo::mutinformation(data[,gene1], data[,gene2], method = "mm")
      miMat[gene2, gene1] <- miMat[gene1, gene2]
    }
  }
  return(miMat)
}

```

Add mutual information to the network. This method is efficient for dense
networks. If the network is sparse, one should calcualte MI for individual 
interactions instead of all pairs.
```{r}
actMi <- calculateMI(data = expData, nbins=8)
      mi <- integer(length = length(network$Source))
      for(int in seq_along(network$Source)){
        mi[int] <- actMi[as.character(network$Source[int]), as.character(network$Target[int])]
      }
      network$Mi <- mi

```


## Helper functions

```{r}
# function to remove two nodes connected to each other only
removeNodePairs <- function(tmpNetwork){
  srcGenes <- unique(tmpNetwork$Source)
  for(i in seq_along(srcGenes)){
    src <- tmpNetwork$Source[which(tmpNetwork$Source == srcGenes[i])]
    tgtGenes <- tmpNetwork$Target[which(tmpNetwork$Target == srcGenes[i])]
    if((length(tgtGenes)==1) & (length(src)==1) ){
      if((tgtGenes==src)){
        print(tgtGenes)
      tmpNetwork <- tmpNetwork[-(which(tmpNetwork$Source == tgtGenes)),]
      tmpNetwork <- tmpNetwork[-(which(tmpNetwork$Target == tgtGenes)),]
      }
}
  }
  return(tmpNetwork)
}
```

## Function to generate the networks

```{r pressure, echo=TRUE}
library(sRACIPE)
library(stringr)
library(roxygen2)


#' @param tmpNetwork A network containing the TF, target and the metric.
#' The metric can be mutual information. It should also have a metric to 
#' identify the sign of the interaction, for example, correlation.
#' In this case the network can have same interaction with different
#' mutual information and/or sign metric as the interaction can come from
#' different sources.
#' @param coreTf List of factors whose nearest neighbors are to be included in 
#' the network. Use all tfs as coreTfs if no core tf is to be used.
#' @param posMi Cutoff for positive mutual information
#' @param negMi Cutoff for negative mutual information
#' @param name generic name for dataset
#' @param simulate whether to simulate the networks using sRACIPE
#' @param allowAllSameSign wether to simulate networks with interactions of
#' one type only
#' @param minCoreTFs minimum number of coreTFs that should be in the network
#' for it to be included in simulations
#' @param minPosInt minimum number of excitatory interactions that should be in the network
#' for it to be included in simulations.
#' @param minNegInt minimum number of inhibitory interactions that should be in the network
#' for it to be included in simulations.
#'
#' 
createNetworkMiValue <- function(tmpNetwork=network,coreTf = coreTf, 
                                 posMi = posMi,
                                 negMi = negMi, name = "A549", 
                                 simulate = FALSE, allowAllSameSign = FALSE,
                                 minCoreTFs = 0, minPosInt=0,minNegInt=0){
  tmpNetwork <- tmpNetwork[union(which(tmpNetwork$Source %in% coreTf),
                                 which(tmpNetwork$Source %in% coreTf)),]
  
  tmpNetwork <- tmpNetwork[(tmpNetwork$Source ),]
  # sort the network based on mutual information
  tmpNetwork <- tmpNetwork[order(tmpNetwork$Mi, decreasing = TRUE),]
  # remove any duplicate interactions
  tmpNetwork <- tmpNetwork[!duplicated(tmpNetwork[c("Source","Target","Type")]),]
  # Select the positive network
  posNetwork <- tmpNetwork[tmpNetwork$Cor >0,]
  posNetwork <- posNetwork[which(posNetwork$Mi > posMi),]
  # Select the negative network
  negNetwork <- tmpNetwork[tmpNetwork$Cor < 0,]
  negNetwork <- negNetwork[which(negNetwork$Mi > negMi),]
  
  # Exit if there are either no positive or negative interaction.
  # Can be removed if all positive or all negative networks are ok
  if(!allowAllSameSign){
  if(dim(posNetwork)[1] == 0 | dim(negNetwork)[1] == 0) return()
    }
# Combine the network
  tmpNetwork <- rbind(posNetwork,negNetwork)
  print("check if there are any duplicate interactions with same or conflicting sign")
  print(sum(duplicated(tmpNetwork[,c("Source","Target")])))
  print(" check if there are any duplicate interactions with same sign")
  print(sum(duplicated(tmpNetwork[,c("Source","Target","Type")])))
  # Remove duplicate interactions
  tmpNetwork <- tmpNetwork[!duplicated(tmpNetwork[,c("Source","Target","Type")]),]
  print("Check if any conflicting interaction is there")
  print(tmpNetwork[duplicated(tmpNetwork[,c("Source","Target")]),])
  print(tmpNetwork[duplicated(tmpNetwork[,c("Source","Target")], fromLast = TRUE),])
# Remove conflicting interactions
  delRows <- anyDuplicated(tmpNetwork[,c("Source","Target")])
  delRows <- c(delRows, anyDuplicated(tmpNetwork[,c("Source","Target")], fromLast = TRUE))
  delRows <- delRows[delRows>0]
  if(length(delRows>0)){ tmpNetwork <- tmpNetwork[-delRows,]}

  # Remove signaling interactions
  # Uncomment for iterative removal
  # interactionRemoved = length(tmpNetwork$Source)
  # while(interactionRemoved>0)
  {
    tmpVar = length(tmpNetwork$Source)
    tmpNetwork <- tmpNetwork[(which(tmpNetwork$Source %in% tmpNetwork$Target )),]
    # if targets only nodes are also to be removed.
    # tmpNetwork <- tmpNetwork[which(tmpNetwork$Target %in% tmpNetwork$Source),]
    # interactionRemoved = tmpVar - length(tmpNetwork$Source)
  }

  tmpNetwork <- tmpNetwork[!duplicated(tmpNetwork),]
  tmpNetwork <- removeNodePairs(tmpNetwork)
  print("Number of interactions")
  print(length(tmpNetwork$Source))
  networkTfs <- union(tmpNetwork$Source,tmpNetwork$Target)
  print("Core Tfs included in the network")
  print(networkTfs[which(networkTfs %in% coreTfGenes)])

  if(length(networkTfs) < minCoreTFs) simulate = FALSE

  if(length(which(tmpNetwork$Type ==2)) <minNegInt) simulate = FALSE
  if(length(which(tmpNetwork$Type ==1)) <minNegInt) simulate = FALSE
  require(igraph)
  circuit <-  tmpNetwork[,c("Source", "Target", "Type")]
  g <- graph_from_data_frame(circuit, directed = TRUE, vertices = NULL)
  if(!igraph::is_connected(g)) simulate = FALSE 


  if(simulate){
    require(sRACIPE)
    tmp <- sracipeSimulate(circuit = tmpNetwork[,c("Source", "Target", "Type")], plotToFile = TRUE,
                           numModels = 2000, plots = FALSE, stepper = "RK4", integrateStepSize = 0.05)
    annotation(tmp) <- paste0(name,"_",as.character(posMi), "_",as.character(negMi),"Interactions")
    # tmp <- sracipePlotData(tmp, plotToFile = T)
    saveRDS(tmp, file = paste0("simAll/",name,"_",as.character(posMi), "_",as.character(negMi),".rds"))
  }
  # return(tmpNetwork)
}

```

## Construct and simulate networks

```{r}
for(posMi in seq(1,0.05,-0.05)){
  for(negMi in seq(0.5,0.05,-0.05)){

    print(paste0("network_",as.character(posMi),"_",as.character(negMi)))
    createNetworkMiValue(tmpNetwork = network,posMi=posMi,negMi = negMi,
                                         name="net", simulate = T)
  }
}

```